---
title: "Demo GLM approach"
author: "A.I."
format:
  html:
    number-sections: true
editor: visual
---

```{r}
#| warning: false

library(PhyloSim)
library(parallel)
library(dplyr)
root <- "~/cyber_synch/" # work from uni bayreuth server
```

To get stabilizing CDD, the first step is to run a binomial GLM with mortality ~ number of conspecific neighbors. With the resulting coefficient of the predictor we can calculate stabilizing CDD according to Hülsmann et al.(2024). This is a short tutorial on how to run the GLM with the simulated data.

## Data preparation

I show, that the same radius for counting conspecific neighbors should be used within one batch to guarantee comparability.
We load one batch (= `runs`) but count conspecific neighbors in two different ways, for didactic purpose. The function `getConNeigh()` does this count. For details on the kernel definition see expKernelDemo.qmd XXXX.  

`RadiusMax` counts conspecific neighbors within a fixed spatial kernel of radius = 5, regardless of the actual DD effect radius in each run. For example, if both nDD (`nDensityCut`) and pDD (`pDensityCut`) equal 1, only the four nearest neighbors in a von Neumann neighborhood have a DD effect on the focal cell. Nevertheless, using `radius = 5` in `getConNeigh()` counts all conspecifics within a radius of 5 corresponding to 80 neighbors.

In contrast, `RadiusNULL` uses `radius = NULL`, which sets the neighborhood radius to the maximum of nDensityCut and pDensityCut for each run. For instance, if a run has nDensityCut = 2 and pDensityCut = 4, conspecifics are counted within a radius = 4. Thus, the largest DensityCut in a run determines the effective neighborhood for counting conspecifics. 

We load the raw runs and apply both described methods.
```{r}
#| eval: false

runs <- readRDS(paste0(root, "/local/runs/mstr/20250903/ranRuns_i.rds"))

RadiusMax <- getConNeigh(runs, radius = 5)
RadiusNULL <- getConNeigh(runs, radius = NULL)

TabMax <- getMatToTab(RadiusMax, detailedParams = T)
TabNULL <-getMatToTab(RadiusNULL, detailedParams = T) 
```

```{r}
#| include: false

TabMax <- readRDS(paste0(root, "/local/runs/mstr/20250903/ranTabMax_i.rds"))
TabNULL <- readRDS(paste0(root, "/local/runs/mstr/20250903/ranTab_i.rds"))
```

Very rare species are filtered out, to increase statistical robustness and to ensures GLM runs without error.
Then binomial GLM is fitted and the estimates extracted.
```{r}
#| eval: false

cl <- makeCluster(25)
clusterExport(cl, c("getMatToTab", "TabNULL"))

clusterEvalQ(cl, {
  library(dplyr)
})

# ------------ sort species wise data ------------
tabNULLs <- parLapply(cl, TabNULL, function(x) {
  x %>%
    filter(abund > 100) %>%
    mutate(specIdCen = paste0(specId, census)) %>%
    select(-indId)
})
# ------------ run glm ------------
FitNULL <- parLapply(cl, tabNULLs, function(x) {
    mod <- glm(mortNextGen ~ con, data = x, family = binomial())
    mort0 <- plogis(coef(mod)[1])
    mort1 <- plogis(coef(mod)[1] + coef(mod)[2])
    res <- data.frame(
      mort_change = mort1 - mort0,
      coef = coef(mod)[2]
    )
  return(res)
})
stopCluster(cl)

# ----------------------------------------------
# ----------------------------------------------
# ----------------------------------------------

cl <- makeCluster(25)
clusterExport(cl, c("getMatToTab", "TabMax"))

clusterEvalQ(cl, {
  library(dplyr)
})


# ------------ sort species wise data ------------
TabMaxs <- parLapply(cl, TabMax, function(x) {
  x %>%
    filter(abund > 100) %>%
    mutate(specIdCen = paste0(specId, census)) %>%
    select(-indId)
})

# ------------ run glm ------------
FitMax <- parLapply(cl, TabMaxs, function(x) {
    mod <- glm(mortNextGen ~ con, data = x, family = binomial())
    mort0 <- plogis(coef(mod)[1])
    mort1 <- plogis(coef(mod)[1] + coef(mod)[2])
    res <- data.frame(
      mort_change = mort1 - mort0,
      coef = coef(mod)[2]
    )
  return(res)
})
stopCluster(cl)
```
# Binomial GLM

Data is restructured
```{r}
#| eval: false

resMax <- do.call(rbind.data.frame, FitMax)
rownames(resMax) <- str_remove(pattern = "\\.\\d$", string = rownames(resMax))

resNULL <- do.call(rbind.data.frame, FitNULL)
rownames(resNULL) <- str_remove(pattern = "\\.\\d$", string = rownames(resNULL))
```

```{r}
#| include: false

# saveRDS(resMax, paste0(root, "/local/runs/mstr/20250903/ranGlmMax.rds"))
# saveRDS(resNULL, paste0(root, "/local/runs/mstr/20250903/ranGlmNULL.rds"))

resMax  <- readRDS(paste0(root, "/local/runs/mstr/20250903/ranGlmMax.rds"))
resNULL <- readRDS(paste0(root, "/local/runs/mstr/20250903/ranGlmNULL.rds"))
```

Estimator values (i.e., coefficients) of nb. conspecific neighbors are plotted. 
```{r}
#| fig-height: 3
#| fig-width: 9.5
#| fig-cap: "Estimator of the effect of conspecific neighbors on mortality from a binomial GLM after plogis() backtransformation. Left, with conspecific count for every run within a radius of rMax, whereby rMax is the maximum radius (i.e., parameter = densityCut) within the batch. Middle, the conspecific count for each run within a batch is determined by the maximum radius within each run. Thus, runs have different radii to count conspecific neighbors. Right, both methods compared."


par(mfrow = c(1,3))
plot(resMax$coef, ylim = c(-0.0015,0.0015), ylab = "coefficient")
plot(resNULL$coef, ylim = c(-0.0015,0.0015), ylab = "coefficient")
boxplot(
  list(RadiusMax = resMax$coef, RadiusNULL = resNULL$coef),
  ylim = c(-0.0015, 0.0015),
  ylab = "coeficient"
)
```
This analysis shows that using the maximum radius within a batch gives more reliable coefficients and make the runs comparable within a batch.
