---
title: "PhyloSim Changes"
author: "A.I."
format:
  html:
    number-sections: true
editor: visual
---

```{r}
#| warning: false
library(PhyloSim)
library(parallel)
library(dplyr)
library(tidyverse)
library(lattice)
library(ggplot2)
library(metafor)
library(MASS)
# root <- "~/Uni/Master/MA/" # work from local machine
root <- "~/cyber_synch/" # work from uni bayreuth server
```

**Here, I reduced the base mortality and the fitness independent mortality. This increases the effect of DD on mort change**

# Quantification of stabilazation: Mortality change for + 1 conspecific individual {#sec-mortality-change}

According to Hülsmann et al. 2024 I quantify the effect of stab CDD by - fitting a binomial model - calculate probability of mortality without conspecific neighbors (= mort0) - calculate probability of mortality with *one* conspecific neighbor (= mort1) - calcualte change in probability (mortChange = mort1 - mort0)

## Data preparation

Note on *fitness base mortality ratio* (fbmr): it skips fitness dependent death every fbmr'th step. E.g., fbmr = 10 would cause always death, independent of fitness. In this batch, it is set to 10. In the next, to 3000. We'll see if fbmr changes the mortChange coefficient dramatically. But, first things first.

```{r}
#| eval: false

# load in runs with exp kernel
runsRaw  <- readRDS(paste0(root, "/local/runs/mstr/20250807/runs_ix.rds"))
```

Next, we

-   calculate the numbers of conspecific neighbors
-   asess if an individuls dies in the consequent generation. Therefore, we ran the simulations with sort(c(seq(x,y), seq(x,y)+1)). Again, we assess death after each period in time trough the immediately next generation (e.g, focal generation 1000 -\> individuals dead at 1001 ?)

```{r}
#| eval: false

# get conspecific neighbors and proper naming
runsRaw  <- getConNeigh(runsRaw)
```

```{r}
# make the runs more slim
runs_ix <- lapply(runsRaw, function(x){
  x$Output <- x$Output[101:300]
  return(x)
})
```

Next, we convert the matrix data into tabular data. With the argument detailedParams we include the parameter settings a seperate cols. We save the tabular data.

```{r}
#| eval: false

# convert matrices to tabular data. This is done parallel, as it takes longer
cl <- makeCluster(length(runs_ix))
clusterExport(cl, c("getMatToTab", "runs_ix"))
tab_ix <- parLapply(cl = cl, X = runs_ix, fun = function(x) getMatToTab(x, detailedParams = TRUE))

saveRDS(tab_ix, paste0(root, "local/runs/mstr/20250807/tab_ix.rds"))
```

We only keep important/varying parameters in the naming. Fixed params are excluded.

```{r}
tab_ix <- readRDS(paste0(root, "local/runs/mstr/20250807/tab_ix.rds"))
```

```{r}
namesShort <- names(tab_ix) %>%
  stringr::str_remove("_disp.+") %>% 
  stringr::str_remove("Cut1") %>% 
  stringr::str_remove("Cut1") %>% 
  stringr::str_replace("pdd", "P") %>% 
  stringr::str_replace("ndd", "N") %>% 
  stringr::str_replace_all("Var", "-L")
```

Now, we delete every second generation. Remember, we had to calculate the death in the consequent generation. After doing so, the generation x + 1 is no longer needed and is discarded.

```{r}
# keep only first timespot in census
tab_ixS <- lapply(tab_ix, function(x){
  res <- x %>%
  dplyr::filter(census %% 2 == 0,
          abund > 100) %>%
    mutate(specIdCen = paste0(specId, census)) %>% 
    dplyr::select(-fbmr, -sr, -fao, -disp, -indId)
  return(res)
})

names(tab_ixS) <- namesShort
```

### Metafor analysis: correcting for uncertainty

```{r}
#| eval: false

cl <- makeCluster(length(tab_ixS))

mcS_err <- parLapply(cl, tab_ixS, function(x) {
  specIDs <- unique(x$specIdCen)
  res <- vector("list", length(specIDs))
  
  i <- 1
  for (sID in specIDs) {
    dat <- x[x$specIdCen == sID, ]
    mod <- glm(mortNextGen ~ con, data = dat, family = binomial())
    sfm <- summary(mod)$coefficients
    vc <- vcov(mod)[c("(Intercept)", "con"), c("(Intercept)", "con")]

    mort0 <- plogis(coef(mod)[1])
    mort1 <- plogis(coef(mod)[1] + coef(mod)[2])
    
    res[[i]] <- list(
      specId = sID,
      abund = dat$abund[1],
      mort_change = mort1 - mort0,
      coef = coef(mod)[c(1,2)],
      vcov = vc
    )
    i <- i + 1
  }
  return(res)
})

# Stop the cluster
stopCluster(cl)
```

## compute variance of the marginal effect through a "posterior" simulation

```{r}
#| eval: false

cl <- makeCluster(length(mcS_err))
mcS_err_sim <- parLapply(cl, mcS_err, function(x){
  lapply(x, function(y){
    sim <- MASS::mvrnorm(n = 100, mu = c(y$coef[1], y$coef[2]), Sigma = y$vcov)
    mort0 <- plogis(sim[, 1])
    mort1 <- plogis(sim[, 1] + sim[, 2])
    mort_diff <- mort1 - mort0
    
    return(data.frame(
      abund = y$abund,
      specId = y$specId,
      mean = mean(mort_diff),
      se = sd(mort_diff),
      ci_low = quantile(mort_diff, 0.025),
      ci_high = quantile(mort_diff, 0.975)
    ))
  })
})
# Stop the cluster
stopCluster(cl)
```

## unlist inner lists and add log abund

```{r}
#| eval: false

m4 <- lapply(mcS_err_sim, function(group) {
  do.call(rbind, group)
})

m4 <- lapply(m4, function(group) {
  row.names(group) <- NULL
  group <- group %>% 
    mutate(log_N = log(abund))
  return(group)
})
```

## fitting model

```{r}
#| eval: false

dat_meta <- lapply(m4, function(x) {
  escalc(measure = "GEN", yi = mean, sei = se, slab = specId, data = x)
})
```

```{r}
#| eval: false

# # Detect available cores and create a cluster
# cl <- makeCluster(length(dat_meta))
# 
# # Export needed objects and packages to the workers
# clusterExport(cl, varlist = c("dat_meta"), envir = environment())
# 
# # Run in parallel
# metamod <- parLapply(cl, dat_meta, function(x) {
#   metafor::rma(
#     yi = yi,
#     vi = vi,
#     mods = ~ log_N,
#     method = "REML",
#     data = x
#   )
# })
# 
# # Stop cluster
# stopCluster(cl)
# 
# names(metamod) <- namesShort
```

```{r}
cl <- makeCluster(length(dat_meta))
clusterExport(cl, varlist = c("dat_meta"), envir = environment())

metamod <- parLapply(cl, dat_meta, function(x) {
  tryCatch({
    metafor::rma(
      yi = yi,
      vi = vi,
      mods = ~ log_N,
      method = "REML",
      data = x
    )
  }, error = function(e) {
    return(NULL)  # Return NULL for failed models
  })
})

stopCluster(cl)

# Remove NULL elements and keep names aligned
failed_indices <- sapply(metamod, is.null)
metamod <- metamod[!failed_indices]
names(metamod) <- namesShort[!failed_indices]

# Report which models failed
if(any(failed_indices)) {
  cat("Failed models:", paste(namesShort[failed_indices], collapse = ", "), "\n")
}
```

```{r}

# deleting one scenario

saveRDS(dat_meta, paste0(root, "local/runs/mstr/20250807/dat_metaix.rds"))
saveRDS(metamod, paste0(root, "local/runs/mstr/20250807/metamodix.rds"))


# metamod <- readRDS(paste0(root, "local/runs/mstr/20250807/metamodiii.rds"))
# dat_meta <- readRDS(paste0(root, "local/runs/mstr/20250807/dat_metaiii.rds"))
```

## predictions

### log_N

```{r}

pred <- lapply(dat_meta, function(x){
  expand_grid(log_N = seq(min(x$log_N, na.rm = TRUE),
                          max(x$log_N, na.rm = TRUE),
                          length.out = 1000))
})
pred <- lapply(pred, function(x){
  x$abund <- exp(x$log_N)
  return(x)
})
# Bind predictions to dataframe
pred <- lapply(seq_along(pred), function(i){
  x <- pred[[i]]
  y <- metamod[[i]]
  
  return(cbind(x, predict(object = y, newmods = x$log_N)))
})

names(pred) <- namesShort
```

```{r}
# Function to extract parameters from names (ignoring Cut3)
extract_params <- function(name) {
  # Remove Cut3 from the name first
  name_clean <- gsub("Cut3_", "_", name)
  
  # Extract P, N, and L values from names like "P0.5-L20_N0-L20"
  p_part <- sub("_.*", "", name_clean)  # Get part before underscore
  n_part <- sub(".*_", "", name_clean)  # Get part after underscore
  
  # Extract P and PL values
  p_val <- as.numeric(sub("-.*", "", sub("P", "", p_part)))
  pl_val <- as.numeric(sub(".*-L", "", p_part))
  
  # Extract N and NL values  
  n_val <- as.numeric(sub("-.*", "", sub("N", "", n_part)))
  nl_val <- as.numeric(sub(".*-L", "", n_part))
  
  return(data.frame(P = p_val, PL = pl_val, N = n_val, NL = nl_val))
}

# Convert pred list to dataframe with parameters
pred_df <- do.call(rbind, lapply(seq_along(pred), function(i) {
  df <- pred[[i]]
  params <- extract_params(names(pred)[i])
  df$scenario <- names(pred)[i]
  df$P <- params$P
  df$PL <- params$PL
  df$N <- params$N
  df$NL <- params$NL
  return(df)
}))

# Assume predict() returns fitted values and confidence intervals
# Adjust column names based on your actual metafor predict output
colnames(pred_df)[colnames(pred_df) == "pred"] <- "fitted"
if("ci.lb" %in% colnames(pred_df)) {
  pred_df$ci_lower <- pred_df$ci.lb
  pred_df$ci_upper <- pred_df$ci.ub
} 

# Create the null reference (P0, N0)
null_data <- pred_df[pred_df$P == 0 & pred_df$N == 0, ]

# Create individual plots for each scenario combination

# Get all unique P and N combinations
p_values <- sort(unique(pred_df$P))
n_values <- sort(unique(pred_df$N))

plots_list <- list()
plot_counter <- 1

# Loop through each P value
for (p_val in p_values) {
  # Loop through each N value  
  for (n_val in n_values) {
    
    # Skip null scenario as separate plot
    if (p_val == 0 & n_val == 0) next
    
    # Get data for this specific P,N combination
    plot_data <- pred_df[pred_df$P == p_val & pred_df$N == n_val, ]
    
    # Skip if no data
    if (nrow(plot_data) == 0) next
    
    # Create color/fill variable based on lambda values
    if (p_val == 0) {
      # Only N varies, color by NL
      plot_data$group_var <- paste0("NL", plot_data$NL)
      legend_title <- "N Lambda"
    } else if (n_val == 0) {
      # Only P varies, color by PL  
      plot_data$group_var <- paste0("PL", plot_data$PL)
      legend_title <- "P Lambda"
    } else {
      # Both P and N vary, color by combination
      plot_data$group_var <- paste0("PL", plot_data$PL, "_NL", plot_data$NL)
      legend_title <- "PL_NL"
    }
    
    # Count number of unique groups (lines) that will be plotted
    n_groups <- length(unique(plot_data$group_var))
    show_errors <- n_groups <= 4  # Show errors only if 4 or fewer lines
    
    # Create base plot
    p <- ggplot()
    
    # Add null reference (always with errors for reference)
    p <- p +
      geom_ribbon(data = null_data, aes(x = abund, ymin = ci_lower, ymax = ci_upper), 
                  alpha = 0.2, fill = "gray50") +
      geom_line(data = null_data, aes(x = abund, y = fitted), 
                color = "black", linetype = "dashed", size = 1.2, alpha = 0.8)
    
    # Add treatment lines conditionally with/without errors
    if (show_errors) {
      # Include error ribbons when 4 or fewer lines
      p <- p +
        geom_ribbon(data = plot_data, aes(x = abund, ymin = ci_lower, ymax = ci_upper, 
                                         fill = group_var), alpha = 0.3) +
        geom_line(data = plot_data, aes(x = abund, y = fitted, color = group_var), 
                  size = 1.2)
    } else {
      # Only lines when more than 4 groups (no error ribbons)
      p <- p +
        geom_line(data = plot_data, aes(x = abund, y = fitted, color = group_var), 
                  size = 1.2)
    }
    
    # Add labels and theme
    p <- p +
      scale_x_log10() +
      labs(title = paste0("P = ", p_val, ", N = ", n_val, " (", n_groups, " scenarios)"), 
           x = "Abundance per Census", y = "Mortality Change",
           color = legend_title, fill = legend_title) +
      theme_minimal() +
      theme(legend.position = "bottom",
            plot.title = element_text(size = 14, hjust = 0.5))
    
    # Store plot
    plots_list[[plot_counter]] <- p
    names(plots_list)[plot_counter] <- paste0("P", p_val, "_N", n_val)
    plot_counter <- plot_counter + 1
  }
}

# Create separate null plot (always with errors)
null_plot <- ggplot() +
  geom_ribbon(data = null_data, aes(x = abund, ymin = ci_lower, ymax = ci_upper), 
              alpha = 0.3, fill = "black") +
  geom_line(data = null_data, aes(x = abund, y = fitted), 
            color = "black", size = 1.5) +
  scale_x_log10() +
  labs(title = "Null Scenario: P = 0, N = 0", 
       x = "Abundance per Census", y = "Mortality Change") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, hjust = 0.5))

# Add null plot to list
plots_list[["null"]] <- null_plot

# Apply consistent scales to all plots
x_range <- range(pred_df$abund[pred_df$abund > 0 & is.finite(pred_df$abund)], na.rm = TRUE)
y_range <- range(c(pred_df$ci_lower, pred_df$ci_upper), na.rm = TRUE)

common_scales <- list(
  scale_x_log10(limits = x_range, expand = expansion(mult = 0.02)),
  coord_cartesian(ylim = y_range)
)

# Apply scales to all plots
plots_list <- lapply(plots_list, function(p) p + common_scales)

# Print summary information
cat("Total number of plots:", length(plots_list), "\n")
cat("Plot names:", names(plots_list), "\n")

# Count plots with/without error ribbons
plots_with_errors <- 0
plots_without_errors <- 0

for (i in seq_along(plots_list)) {
  if (names(plots_list)[i] != "null") {
    # Extract P and N values from plot name to get the data
    plot_name <- names(plots_list)[i]
    p_val <- as.numeric(sub("P([0-9.]+)_N.*", "\\1", plot_name))
    n_val <- as.numeric(sub(".*_N([0-9.]+)", "\\1", plot_name))
    
    plot_data <- pred_df[pred_df$P == p_val & pred_df$N == n_val, ]
    if (p_val == 0) {
      group_var <- paste0("NL", plot_data$NL)
    } else if (n_val == 0) {
      group_var <- paste0("PL", plot_data$PL)
    } else {
      group_var <- paste0("PL", plot_data$PL, "_NL", plot_data$NL)
    }
    n_groups <- length(unique(group_var))
    
    if (n_groups <= 4) {
      plots_with_errors <- plots_with_errors + 1
    } else {
      plots_without_errors <- plots_without_errors + 1
    }
  }
}

cat("Plots with error ribbons (≤4 lines):", plots_with_errors, "\n")
cat("Plots without error ribbons (>4 lines):", plots_without_errors, "\n\n")

# Display all plots
for (i in seq_along(plots_list)) {
  cat("Displaying plot:", names(plots_list)[i], "\n")
  print(plots_list[[i]])
}
```

## save PDF

```{r}
pdf(paste0(root, "local/figures/mstr/20250807/metafor/corr_128NoDc.pdf"), width = 8, height = 6, onefile = TRUE, useDingbats = FALSE)
lapply(plots_list, print)   # each print() → new page
dev.off()
```

```{r}
image(runs_vi$pdd0Var20Cut1_ndd0Var20Cut1_disp1_sr2_fbmr3000_faoM$Output$`112501`$specMat, col = rainbow(100))
image(runs_vi$pdd0Var20Cut1_ndd1Var20Cut1_disp1_sr2_fbmr3000_faoM$Output$`112501`$specMat, col = rainbow(100))
image(runs_vi$pdd1Var20Cut1_ndd0Var20Cut1_disp1_sr2_fbmr3000_faoM$Output$`112501`$specMat, col = rainbow(100))
```
